The data this week is quite large. [Find it here](https://fileshare.theinformationlab.co.uk/index.php/s/i032l95DDKrDuBj).
```{r setup}
library(data.table)
library(forcats)
library(extrafont)
library(gridExtra)
library(lubridate)
library(stringr)
library(tidyverse)
options(scipen = 999)

setwd("../Desktop")

get_file <- "RDW_Passenger_Cars.csv"
```

This file is humongous (~9M records) so I read in a sample before reading in the whole thing. ```fread()``` in ```library(data.table)``` makes this much faster than ```read.csv()```
```{r raw_data_query}
data_sample <- 
  read.csv(get_file, nrows = 10000, stringsAsFactors = F)

raw_data <- 
  fread(get_file, select = c(3, 6, 8, 21), #nrows = 500,
        check.names = T, data.table = F) %>% 
  mutate(year.admission = dmy(date.first.admission) %>% year(.),
         date.ascription = dmy(date.ascription),
         year.ascription = year(date.ascription),
         week.ascription = week(date.ascription))
```

Because I need this to be manageable, I use ```fct_lump()``` from ```library(forcats)``` to lump together the largest groups for brand and setup
```{r data_rollup_query}
data_rollup <-
  raw_data %>% 
  mutate(brand = fct_lump(brand, n = 10),
         setup = fct_lump(setup, n = 10)) %>% 
#  group_by(brand) %>% summarise(n = n()) %>% arrange(-n)
  group_by(year.ascription, year.admission, brand, setup) %>% 
  summarise(N = n()) %>%
  ungroup()

write.csv(data_rollup, "2017_19.csv", row.names = F)

rm(raw_data)
```

This decade query cuts the data into 5 year intervals using a custom ```round_down()``` function. The intent of the analysis is to look at the proportion of vehicles registered by year that vehicle was created. I expect that there are much fewer vehicles registered in year 9 than in year 2. I also expect that the highest point for registration will be the year the vehicle becomes available.
```{r decade_query}
break_by <- 10

round_down <-
  function(x){
    x - (x %% break_by)
  }

decade <- 
  data_rollup %>% 
  filter(year.ascription >= 1970) %>% 
  mutate(decade = round_down(year.admission),
         decade = paste0(decade, " - ", decade + break_by - 1),
         decade = ifelse(year.admission >= 2010, "2010 - 2015", decade)
           #cut(year.admission, breaks = seq(1949,2020, 5))
         ) %>% 
  group_by(year.admission, decade, year.ascription) %>% 
  summarise(n = sum(N)) %>% 
  group_by(year.ascription) %>% #####this one:  year.ascription 
  mutate(pct = n / sum(n)) %>% 
  ungroup() %>% 
  mutate(year_number = year.admission %% break_by,
         years_in_market = year.ascription - year.admission + 1,
         pct = round(pct, 4) * 100) %>% 
  filter(years_in_market <= 20,
         year.admission >= 1970,
         year.admission <= 2015)
```

To label my line charts, I need to identify the vehicle year within each 5-year intervals that was most registered as well as the second and third highest peaks. 
```{r peaks}
#peaks
peak_1 <-
  decade %>% 
  filter(years_in_market == 1) %>% 
  group_by(decade) %>%
  filter(pct == max(pct)) %>% 
  ungroup() %>% 
  mutate(label = paste0(year.admission),
         peak = 1)

peak_2 <-
  decade %>% 
  filter(years_in_market > 1) %>% 
  group_by(decade) %>% #### decade
  filter(pct == max(pct)) %>% 
  ungroup() %>% 
  mutate(label = years_in_market, 
         peak = 2)

peak_3 <-
  decade %>%
  filter(years_in_market != 1) %>% 
  anti_join(select(peak_2, decade, years_in_market)) %>%
  anti_join(select(peak_2, decade, year.admission)) %>% 
  group_by(decade) %>%
  filter(pct == max(pct)) %>% 
  ungroup() %>% 
  mutate(label = years_in_market,
         peak = 3)

peaks <-
  rbind(peak_1, peak_2, peak_3) %>% 
  mutate(size = ifelse(peak == 1, 6, 4),
         x = years_in_market,
         y = pct,
         xend = ifelse(peak == 1, years_in_market + 1, years_in_market),
         yend = ifelse(peak == 1, pct, 
                ifelse(peak == 2, pct + 10, pct + 5))
         )
```

I also need a table for my heatmap.
```{r}
peak_each_year <-
  decade %>% 
  filter(years_in_market > 1) %>% 
  group_by(year.admission) %>%
  arrange(-pct) %>% 
  slice(1) %>% 
  ungroup()
```

Separating out the titles here allows me to edit them more efficiently. I am using the ```str_wrap()``` command from ```library(stringr)``` 
```{r plot_titles}
plot_title_subtitle <-
  "MakeoverMonday - Vehicle Registration in the Netherlands: Given the data, these charts show the percentage of vehicles registered each year by the year the vehicle was made. Each line in the charts, and each row on the heatmap, represent the year the vehicles were made, and the x-axis is the length of time the vehicles have been in the market. Changes in the 1990s may be attributed to increased leasing rates, although more research is needed." %>% 
  str_wrap(170)

plot_lines_title <-
  "There are 10 lines in each panel. Each line represents the proportion of registrations for vehicles released in that year over time. The large number at the beginning of each panel represents the production year that had the highest proportion of registrations in the first year of its release. The dashed line indicates the next largest peak." %>% 
  str_wrap(120)

plot_heatmap_title <-
  "The dot indicates the largest peak after the year of release.\n\n."

```

Here I create 3 plots.
* A small plot for the title & subtitle
* Line charts
* Heatmap

I then use ```grid.arrange()``` by ```library(gridExtra)``` to piece them all together
```{r plot_create}
font_family <- "xkcd"

font_import(pattern = "font_family", prompt = F)

fonttable()
  
  
plot_title <-{
  ggplot()+
  ggtitle(label = "Cars released after 1990 have second wind 4-5 years after entering the market",
          subtitle = plot_title_subtitle) +
  coord_fixed(ratio = .01) +
  theme(text = element_text(family = font_family, hjust = 1),
        plot.title = element_text(size = 20, color = "grey30", face = "bold"),
        plot.subtitle = element_text(size = 14, color = "grey50"))
}
plot_title

plot_lines <-{
  ggplot(decade, aes(x = years_in_market, y = pct)) +
  facet_wrap(~ decade, nrow = 2) +
  scale_x_continuous(breaks = seq(0, 20, 5)) +
  geom_line(aes(group = year.admission, color = year_number), 
            alpha = .3, size = 1, alpha = .6,
            position = position_dodge(width = 0.7)) +
  scale_color_gradient(#low = "grey70", high = "navyblue", 
                       name = "Year Released \nin Decade",
                       breaks = c(1, 5, 10, 11)) +
  ylim(0, 48) +
  labs(x = "Years in Market", 
       y = "Percent of Vehicles Registered",
       caption = plot_lines_title) +
  geom_vline(data = filter(peaks, peak == 2), 
             aes(xintercept = x),
             color = "navyblue", linetype = 2, alpha = .5) +
  geom_segment(data = filter(peaks, peak == 1), 
               aes(xend = xend, yend = yend),
               color = "grey50") +
  geom_text(data = filter(peaks, peak == 1), 
            aes(x = xend, y = yend, label = label),
            vjust = "center", hjust = "left", size = 6, color = "grey50", family = font_family) +
  theme(text = element_text(size = 12, family = font_family),
        panel.background = element_rect(fill = "grey95", color = "grey50"),
        strip.text.x = element_text(size = 16),
        plot.caption = element_text(hjust = 0, size = 12, color = "grey30"),
        legend.title = element_text(color = "grey30"), 
        legend.position = c(0.85, 0.1),
        legend.direction = "horizontal", 
        panel.grid.minor.y = element_blank()
        )
}

plot_lines


plot_heatmap <-{
  decade %>% 
  mutate(pct = ifelse(pct > 30, 30, pct),
         year.admission = fct_rev(factor(year.admission))) %>% 
  ggplot(aes(x = years_in_market, y = as.factor(year.admission), fill = pct)) +
    facet_grid(decade ~ ., scales = "free_y", space = "free_y") +
    geom_tile(color = "grey75") +
    geom_point(data = peak_each_year, color = "grey40") +
#    scale_y_reverse(breaks = seq(2020, 1970, -5)) + 
    scale_fill_gradient2(low = "white", high = "navyblue", midpoint = .01, 
                         name = "Percent of \nVehicles \nRegistered") +
    labs(x = "Years in Market",
         y = "Year Released",
         caption = plot_heatmap_title) +
    theme(text = element_text(size = 12, family = font_family), 
          panel.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank(),
          plot.caption = element_text(hjust = 0,size = 12, color = "grey30"),
          legend.title = element_text(color = "grey30"), 
          legend.position = c(0.77, 0.1),
          legend.direction = "horizontal"
          )
}
plot_heatmap


grid.arrange(plot_title, plot_lines, plot_heatmap,
             layout_matrix = rbind(c(1,1,1,1,1),
                                   c(2,2,2,3,3),
                                   c(2,2,2,3,3),
                                   c(2,2,2,3,3),
                                   c(2,2,2,3,3)))
```

#Not using
```{r not_using}
ggplot(data_rollup) +
  geom_col(aes(year.ascription, N)) +
  xlim(1975, 2016)

data_sample %>% 
  mutate(var = catalog.price,
         date = dmy(date.ascription),
         year = year(date)) %>%
  filter(year > 2004, year < 2017) %>% 
  select(date.ascription, var, year) %>%
  group_by(year) %>% 
  summarize(var = median(var, na.rm = T)) %>% 
  ungroup()  %>% 
  as_tibble() %>% 
  ggplot(aes(year, var)) +
    geom_col()

data_rollup %>%
  filter(year.admission > 2000) %>% 
  group_by(year.admission, year.ascription, brand) %>% 
  summarise(n = sum(N)) %>% 
  group_by(year.admission, brand) %>% 
  mutate(freq = n / sum(n) * 1000) %>% 
  ungroup() %>% 
  mutate(year.diff = year.ascription - year.admission) %>% 
  ggplot() +
    geom_point(aes(year.admission, year.ascription, 
                   size = n, color = factor(year.admission))) +
    facet_wrap(~brand)
  ```
