---
title: "Internet Usage per 100 people 2005 - 2015"
subtitle: "MakeoverMonday challenge for week 22 2017"
date: "May 30, 2017"
author: "jake riley"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, 
                      cache = T,
                      fig.height = 3.5, fig.width = 4)
```

####Find more information here:

* [MakeoverMonday](http://www.makeovermonday.co.uk/data/)
* [Original Data](https://knoema.com/atlas/topics/Telecommunication/Internet-Users/Internet-users-per-100-people?action=export&gadget=tranking-container)
* [Code on GitHub/rjake](https://github.com/rjake/makeoverMonday/blob/master/2017_22_Internet_Users.Rmd)


```{r setup_workspace}
library(googlesheets)

library(tidyverse)
library(plotly)

options(scipen = 999)

challenge_week <- 22
```

```{r read_data}
#google sheet to read in
  google_key <- 
    gs_key("1XpDdIfrHLaL5Iiu7x8eVAIlDuq83zwUNqJSUsYSAh18")

#get names of worksheets
#  gs_ws_ls(google_key)

#get raw data
  data_raw <- 
    gs_read(google_key, 
            ws = as.character(challenge_week), 
            lookup = T) %>% 
    data.frame()
    
  data_clean <-
    data_raw  %>%
    rename(users_per_100 = internet_users_per_100_people) %>% 
    group_by(country) %>% 
    filter(year >= 2005) %>% 
    filter(year == min(year)|
           year == max(year)) %>%
    filter(!is.na(users_per_100)) %>% 
    filter(n_distinct(year) == 2) %>% 
    ungroup() %>% 
    mutate(year = paste0("year_", year))
```

```{r create_clusters}
  data_spread <-
    data_clean %>% 
    spread(key = year, value = users_per_100) %>% 
    #sample_n(50) %>% 
    mutate(diff = year_2015 - year_2005)

  clusters <- 
    hclust(
      dist(data_spread %>% 
             select(diff, year_2015)
           ))

  data_spread$group <- 
      paste0("cluster ",
             cutree(clusters, 
                    k = 6))
                    #h = mean(clusters$height)))
```

```{r create_outlines}
Regions <- 
  data_spread %>%
  distinct(group)

group_outline <- 
  data.frame(diff = vector("numeric", 0),
             year_2015 = vector("numeric", 0),
             group = vector("character", length = 0))

for(i in 1:nrow(Regions)){
#  i = 2
  group_points <-  
    data_spread %>% 
    filter(group == Regions$group[i]) %>% 
    select(diff, year_2015)
  
  select_hull <- chull(group_points)
  select_hull <- c(select_hull, select_hull[1])
  
  group_outline <- 
    rbind(group_outline,
          group_points[select_hull, ] %>%
    mutate(group = Regions$group[i]))

}
```

###*The median rate of internet users per 100 people has seen a large shift from 2005 to 2015 (n = 185 countries)*
```{r plot_boxplot, fig.height = 3, fig.width = 6}
ggplot(data_clean) +
  geom_boxplot(aes(x = year, 
                   y = users_per_100,
                   group = year)) +
  coord_flip()
```

####These charts use ```plotly```. You can zoom in, exclude values, and hover over points for more information.

###*Growth of internet usage per 100 people (by country) can be grouped into different categories:*
* Low usage rate in 2005 and little growth by 2015 (bottom left)
* High rate in 2005 and small increase in 2015 (bottom right)
* Moderate rate in 2005 and 20-40% growth by 2015 (middle right)
* Low rate in 2005 and 50-70% growth by 2015Large increases to high rates in 2015 (top right)

```{r plot_main}
ggplotly(
  ggplot() +
    geom_polygon(data = group_outline, 
                 aes(x = year_2015, 
                     y = diff, 
                     group = group, 
                     fill = factor(group)), 
                 color = "black",
                 size = 1,
                 alpha = .3) +
    geom_segment(data = data_spread,
                 aes(x = year_2005, xend = year_2015, 
                     y = diff, yend = diff, 
                   color = group),
                 size = .5,
                 alpha = .3) +
    #geom_hline(yintercept = median(data_spread$diff)) +
    #geom_vline(xintercept = median(data_spread$year_2015)) +
    geom_point(data = data_spread,
               aes(x = year_2015, 
                   y = diff, 
                   text = country, 
                   color = group)) +
    geom_text(data = data_spread %>% 
                     filter(country == "United States"),
           aes(x = year_2015 - 10,
               y = diff + 1, label = country)) +
    theme(legend.position="none",
          panel.background = element_rect(fill = NA)) +
    labs(x = "left = 2005, right = 2015 rate per 100 people",
         y = "total difference between 2005 and 2015")
)
```


###*The rate of internet users per 100 people in the United States didn't see much growth between 2005 - 2015*
```{r plot_diag}
ggplotly(  
  data_spread %>% 
  ggplot() +
  geom_segment(aes(x = year_2005, xend = year_2015,
                   y = year_2005, yend = year_2005, 
                   group = country,
                   color = group),
               alpha = .2) +
  geom_point(aes(x = year_2015,
                 y = year_2005, 
                 color = group),
             size = 2) +
  geom_text(data = . %>% filter(country == "United States"),
             aes(x = year_2015 - 13,
                 y = year_2005 + 1, label = country)) +
  theme(panel.background = element_rect(fill = NA))+
  geom_segment(aes(x = 0, y = 0,
                   xend = 100, yend = 100)) +
  coord_equal()+
  xlab("year_2015")
)  
```


.
