```{r setup}
  library(data.table)
  library(ggrepel)
  library(forcats)
  library(lubridate)
  library(tidyverse)
  library(grid)
  library(gridExtra)
  library(stringr)
  options(scipen = 999)

  setwd("C:\\Users\\foxtr\\Desktop")

  get_file <- "RDW_Passenger_Cars.csv"
```

This file is humongous (~9M records) so I read in a sample before reading in the whole thing. ```fread()``` in ```library(data.table)``` makes this much faster than ```read.csv()```
```{r raw_data_query}
data_sample <- 
  read.csv(get_file, nrows = 10000, stringsAsFactors = F)


raw_data <- 
  fread(get_file, select = c(3, 6, 8, 21), #nrows = 500,
        check.names = T, data.table = F) %>% 
  mutate(year.admission = dmy(date.first.admission) %>% year(.),
         date.ascription = dmy(date.ascription),
         year.ascription = year(date.ascription),
         week.ascription = week(date.ascription))
```

Because I need this to be manageable, I use ```fct_lump()``` from ```library(forcats)``` to lump together the largest groups for brand and setup
```{r data_rollup_query}
data_rollup <-
  raw_data %>% 
  mutate(brand = fct_lump(brand, n = 10),
         setup = fct_lump(setup, n = 10)) %>% 
#  group_by(brand) %>% summarise(n = n()) %>% arrange(-n)
  group_by(year.ascription, year.admission, brand, setup) %>% 
  summarise(N = n()) %>%
  ungroup() 

rm(raw_data)
```

This decade query cuts the data into 5 year intervals using a custom ```round_down()``` function. The intent of the analysis is to look at the proportion of vehicles registered by year that vehicle was created. I expect that there are much fewer vehicles registered in year 9 than in year 2. I also expect that the highest point for registration will be the year the vehicle becomes available.
```{r decade_query}
round_down <-
  function(x, round.val){
    x - (x %% round.val)
  }

decade <- 
  data_rollup %>% 
  filter(year.ascription >= 1975) %>% 
  mutate(decade = round_down(year.admission, 5),
         decade = paste0(decade, " - ", decade + 4)
           #cut(year.admission, breaks = seq(1949,2020, 5))
         ) %>% 
  group_by(year.admission, decade, year.ascription) %>% 
  summarise(n = sum(N)) %>% 
  group_by(year.ascription) %>% #####this one:  year.ascription 
  mutate(pct = n / sum(n)) %>% 
  ungroup() %>% 
  mutate(years_in_market = year.ascription - year.admission,
         pct = round(pct, 4) * 100) %>% 
  filter(years_in_market < 20,
         year.admission >= 1975,
         year.admission < 2015)
```

To label my line charts, I need to identify the vehicle year within each 5-year intervals that was most registered as well as the second and third highest peaks. 
```{r peaks}
#peaks
peak_1 <-
  decade %>% 
  filter(years_in_market == 0) %>% 
  group_by(decade) %>%
  filter(pct == max(pct)) %>% 
  ungroup() %>% 
  mutate(label = paste0(year.admission),
         peak = 1)

peak_2 <-
  decade %>% 
  filter(years_in_market > 0) %>% 
  group_by(decade) %>% #### decade
  filter(pct == max(pct)) %>% 
  ungroup() %>% 
  mutate(label = years_in_market, 
         peak = 2)

peak_3 <-
  decade %>%
  filter(years_in_market != 0) %>% 
  anti_join(select(peak_2, decade, years_in_market)) %>%
  anti_join(select(peak_2, decade, year.admission)) %>% 
  group_by(decade) %>%
  filter(pct == max(pct)) %>% 
  ungroup() %>% 
  mutate(label = years_in_market,
         peak = 3)

peaks <-
  rbind(peak_1, peak_2, peak_3) %>% 
  mutate(size = ifelse(peak == 1, 6, 4),
         x = years_in_market,
         y = pct,
         xend = ifelse(peak == 1, years_in_market + 1, years_in_market),
         yend = ifelse(peak == 1, pct, 
                ifelse(peak == 2, pct + 10, pct + 5))
         )
```

I also need a table for my heatmap.
```{r}
peak_each_year <-
  decade %>% 
  filter(years_in_market > 0) %>% 
  group_by(year.admission) %>%
  arrange(-pct) %>% 
  slice(1) %>% 
  ungroup()
```

Separating out the titles here allows me to edit them more efficiently. I am using the ```str_wrap()``` command from ```library(stringr)``` 
```{r plot_titles}
plot_title_subtitle <-
  "#MakeoverMonday - Vehicle Registration in the Netherlands: These charts show the percentage of vehicles registered each year by the year the vehicle was made.The year of the model (each line) are arranged by 5-year intervals and the x-axis is the length of time the vehicles have been in the market. Changes in the 1990s may be attributed to increased leasing rates (more research is needed). In 2000, car registrations took off and cars released in that year made up 40% of the registrations." %>% 
  str_wrap(165)

plot_lines_title <-
  "Proportion of registrations by vehicle year: There are 5 lines in each panel -- one for each year within the panel header. The large dark number at the beginning of each panel represents the year that made up the largest percentage of cars registered within that period. The dashed line indicates the next largest peak after year 0." %>% 
  str_wrap(110)

plot_heatmap_title <-
  "The dot indicates the largest peak after year 0 by year of release."

```

Here I create 3 plots.
* A small plot for the title & subtitle
* Line charts
* Heatmap

I then use ```grid.arrange()``` by ```library(gridExtra)``` to piece them all together
```{r plot_create}
plot_title <-{
  ggplot()+
  ggtitle(label = "Cars released 1990+ have second wind 3-4 years after entering the market",
          subtitle = plot_title_subtitle) +
  coord_fixed(ratio = .01) +
  theme(plot.title = element_text(hjust = 1,size = 20, color = "grey30", face = "bold"),
        plot.subtitle = element_text(hjust = 1, size = 14, color = "grey50"))
}

plot_lines <-{
  ggplot(decade, aes(x = years_in_market, y = pct)) +
  facet_wrap(~ decade, nrow = 3) +
  scale_x_continuous(breaks = seq(0, 20, 4)) +
  geom_line(aes(group = year.admission), 
            color = "navyblue", alpha = .3, size = 1, alpha = .6,
            position = position_dodge(width=0.5)) +
  ylim(0, 45) +
  labs(x = "Years in Market", 
       y = "% of Cars Registered") +
  geom_vline(data = filter(peaks, peak == 2), 
             aes(xintercept = x),
             color = "navyblue", linetype = 2, alpha = .5) +
  geom_segment(data = filter(peaks, peak == 1), 
               aes(xend = xend, yend = yend),
               color = "grey50") +
  geom_text(data = filter(peaks, peak == 1), 
            aes(x = xend, y = yend, label = label, 
                    size = size, color = (peak < 3)),
            vjust = "bottom", hjust = "left") +
  scale_size(range = c(5,6), guide = F) +
  scale_color_manual(values = c("grey50", "grey20"), guide = F) +
  ggtitle(label = plot_lines_title) +
  theme(panel.background = element_rect(fill = "grey95", color = "grey50"),
        plot.title = element_text(hjust = 0, size = 12, color = "grey30"),
        text = element_text(size = 16))
}
plot_lines

plot_heatmap <-{
  decade %>% 
  mutate(pct = ifelse(pct > 30, 30, pct)) %>% 
  ggplot(aes(x = years_in_market, y = year.admission, fill = pct)) +
    facet_grid(decade ~ ., scales = "free_y") +
    geom_tile(color = "grey75") +
    geom_point(data = peak_each_year, color = "grey40") +
    scale_y_reverse(breaks = seq(2015, 1975, -5)) +
    scale_fill_gradient2(low = "white", high = "navyblue", midpoint = .01) +
    labs(y = "year_admission") +
    ggtitle(label = plot_heatmap_title) +
    theme(panel.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank(),
          plot.title = element_text(hjust = 0,size = 12, color = "grey30"))
}
plot_heatmap

grid.arrange(plot_title, plot_lines, plot_heatmap,
             layout_matrix = rbind(c(1,1,1,1,1),
                                   c(2,2,2,3,3),
                                   c(2,2,2,3,3),
                                   c(2,2,2,3,3),
                                   c(2,2,2,3,3)))
```

#Not using
```{r not_using}
ggplot(data_rollup) +
  geom_col(aes(year.ascription, N)) +
  xlim(1975, 2016)

data_sample %>% 
  mutate(var = catalog.price,
         date = dmy(date.ascription),
         year = year(date)) %>%
  filter(year > 2004, year < 2017) %>% 
  select(date.ascription, var, year) %>%
  group_by(year) %>% 
  summarize(var = median(var, na.rm = T)) %>% 
  ungroup()  %>% 
  as_tibble() %>% 
  ggplot(aes(year, var)) +
    geom_col()

data_rollup %>%
  filter(year.admission > 2000) %>% 
  group_by(year.admission, year.ascription, brand) %>% 
  summarise(n = sum(N)) %>% 
  group_by(year.admission, brand) %>% 
  mutate(freq = n / sum(n) * 1000) %>% 
  ungroup() %>% 
  mutate(year.diff = year.ascription - year.admission) %>% 
  ggplot() +
    geom_point(aes(year.admission, year.ascription, 
                   size = n, color = factor(year.admission))) +
    facet_wrap(~brand)
  ```
