---
title: "Toilets in India"
subtitle: "MakeoverMonday challenge for week 32 2017"
date: "August 9, 2017"
author: "jake riley"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F,
                      cache = T#,
                      #fig.height = 3.5, fig.width = 4
)
```

####Find more information here:

* [MakeoverMonday](http://www.makeovermonday.co.uk/data/)
* [Original Data](http://www.asercentre.org/p/289.html)
* [Code on GitHub/rjake](https://github.com/rjake/makeoverMonday/blob/master/2017_32_Toilets_in_India.Rmd)


Making the .gif was no easy task. You will need ImageMagic (zip + installer) and ffmpeg. I recommend these two articles:
  
  * [https://rpubs.com/omicsdata/gganimate](https://rpubs.com/omicsdata/gganimate)
* [https://github.com/dgrtwo/gganimate](https://github.com/dgrtwo/gganimate)

```{r setup_workspace, echo = T}
#library(DT)
#library(forcats)
library(gganimate) #may need: devtools::install_github("dgrtwo/gganimate")
library(animation)
library(googlesheets)
library(ggmap)
library(knitr)
library(raster) #get India's boundary coordinates
library(rgeos) #to simplify file
library(tidyverse)
library(plotly)
library(stringr)
library(tweenr)
library(ggrepel)
```

```{r}
#You may need this if R can't find the convert.exe file
Sys.setenv(PATH = paste("C:/Program Files/ImageMagick/bin", Sys.getenv("PATH"), sep = ";"))

options(scipen = 999)

challenge_week <- "32"
```

```{r read_data}
#google sheet to read in
  google_key <- 
    gs_key("1XpDdIfrHLaL5Iiu7x8eVAIlDuq83zwUNqJSUsYSAh18")

#get data from googlesheets
  data_raw <- 
    gs_read(google_key, 
            ws = as.character(challenge_week), 
            lookup = T, check.names = F) %>% 
    data.frame()
```

#Download and simplify shapefiles
```{r}
#get shapefile
  india_get <- getData('GADM', country='IND', level=1)

#simplify geometry
  india <-
    gSimplify(india_get, tol = .1) %>%
    fortify() %>% 
    filter(lat > 15 | long < 90) %>% 
    mutate(type = "states")

#india <-  map_data("world", region = "India")
  
#plot it out
  ggplot(india) +
    geom_polygon(aes(long, lat, group = group), 
                 fill = "grey60", color = "white") +
    coord_map() +
    geom_point(aes(X, Y), data = data_raw, size = 3)
```

#chart for plot
```{r}
#function to set slope chart on top of map
  range_new <- 
    function(x, newMin, newMax){
      (x - min(x))/(max(x)-min(x)) * (newMax - newMin) + newMin 
    }

#plot of state centroids
  points_state <-
    data_raw %>%
    filter(Category != "Usable Girls Toilet",
           Year != 2014) %>% 
    rename(Pct_Usable_Toilets = X..Usable.Toilets) %>% 
    mutate(State = gsub(" \\(New)", "", State),
           Type = "original",
           Label = "") %>% 
    group_by(State) %>%
    arrange(Year) %>% 
    mutate(Improved = 
             lead(Pct_Usable_Toilets) > Pct_Usable_Toilets,
           Improved = 
             ifelse(is.na(Improved), 
                    lag(Improved, default = T), Improved)) %>% 
    ungroup()

#plot slope chart
  points_slope <-
    points_state %>% 
    group_by(Year) %>% 
    mutate(Extreme = 
             (Pct_Usable_Toilets == max(Pct_Usable_Toilets)) | 
             (Pct_Usable_Toilets == min(Pct_Usable_Toilets))) %>% 
    ungroup() %>% 
    mutate(Label = 
             ifelse(Extreme == T,
                    paste0(
                      round(Pct_Usable_Toilets*100, 0), "%"),
                    NA),
           Y = range_new(Pct_Usable_Toilets, 10, 35),
           X = ifelse(Year == 2010, 75, 90),
           Type = "slope") %>% 
    dplyr::select(one_of(colnames(points_state)))

#bring together
  points_final <-
    rbind(points_slope, points_state)

```

```{r}
#sample of overlay
  ggplot(india) +
    geom_polygon(aes(long, lat, group = group), 
                 data = india, fill = "white", color = "white") +
    coord_fixed() +
    geom_point(aes(X, Y, color = Improved), data = points_slope, size = 3) +
    geom_path(aes(X, Y, group = State, color = Improved), data = points_slope)
```


#Animation
###Animate plot to state centroids
```{r fig.show='animate'}  
#to cutoff
  points_LT_60 <-
    points_slope %>% 
    filter(Year == 2016 & Pct_Usable_Toilets < .6)

#source for frames
  points_sources <- 
    points_final %>%
    dplyr::select(X, Y, Pct_Usable_Toilets, Year,
                  Type, Improved, State, Label) %>% 
    mutate(Type = factor(Type),
           State = factor(State),
           Label = factor(Label),
           Improved = factor(Improved)) %>% 
    group_by(Type) %>%
    nest()

 # points_sources

#create frames with tween_states (tweenr)  
  points_data <- 
    tween_states(points_sources$data, tweenlength = 1,
                 statelength = 0.5, ease = "sine-out", 
                 nframe = 10) %>%
    as.tibble() %>% 
    mutate(Improved = as.logical(Improved),
           Improved = ifelse(.frame <= 3, Improved,
                      !State %in% points_LT_60$State))
  
  #reverse order for animation
    points_data <-
      points_data %>% 
      mutate(.frame = 12 - .frame + 11) %>% 
      rbind(points_data)

p <- 
  ggplot() +
  geom_point(aes(X, Y, color = Improved, frame = .frame), 
             data = points_data, size = 3) +
  geom_path(aes(X, Y, group = State, color = Improved, frame = .frame), 
            data = points_data) +
  geom_text(aes(X, Y, group = State, label = Label, frame = .frame), 
            data = filter(points_data, !is.na(Label)))

animation::ani.options(interval = c(1/2))
gganimate(p, title_frame = FALSE)
```


###Text
```{r fig.show='animate'}
make_labels <-
  function(frame, label, x, y, size){
    tibble(.frame = frame) %>% 
      mutate(label = label, X = x, Y = y, size = size)
  }

points_text <-
  make_labels(c(1:4), "Growth by State", 65, 38, 8) %>% 
  
  rbind(make_labels(c(1:4), "2010", 72,   5, 8)) %>% 
  rbind(make_labels(c(1:12), "2016", 87, 5, 8)) %>%
  
  rbind(make_labels(c(1:4), "24%", 73, 9, 5)) %>% 
  rbind(make_labels(c(1:4), "68%", 73, 30, 5)) %>% 
  rbind(make_labels(c(1:4), "86% Haryana", 91, 35, 5)) %>%
  rbind(make_labels(c(2:4), "40% Mizoram", 91, 16, 5)) %>% 
  
  rbind(make_labels(3, "below 60%", 91, 24, 6))%>% 
  mutate(X = as.numeric(X),
         Y = as.numeric(Y),
         .frame = as.integer(.frame),
         color = !label %in% c("below 60%", "40% Mizoram")) %>% 
  full_join(tibble(.frame = c(1:22)))
  
p <-
  ggplot() +
  geom_text(aes(X, Y, label = label, 
                frame = .frame, size = size, 
                color = color, alpha = size/8), 
            filter(points_text, .frame < 8), 
            hjust = 0) +
  xlim(62, 100) +
  ylim(0, 40) +
  scale_alpha(range = c(.7, 1)) +
  scale_radius(range = c(5, 8)) +
  geom_point(aes(X, Y), points_slope) +
  coord_map()

animation::ani.options(interval = c(1/2))
gganimate(p, title_frame = FALSE)
```



###Animate corners to map of India
```{r}
#final coordinates
  india_outline <-
    india %>%
    mutate(lat = ifelse(lat > 25, max(lat) + 5, min(lat) - 5),
           long = ifelse(long > 78, max(long) + 5, min(long) - 5),
           type = "corner") %>% 
    rbind(india)

#plot of points
  ggplot(india_outline, aes(frame = type)) +
    geom_path(aes(long, lat, color = id, 
                  group = paste0(group,00,order)), 
              alpha = .05, data = india_outline) +
    geom_polygon(aes(long, lat, group = group, fill = id), alpha = .2, 
                 data = india)+#, fill = "white", color = "white") +
    coord_map() +
    guides(fill = F, color = F) +
    geom_point(aes(long, lat, color = id), 
               data = india_outline, size = .8) +
    theme_void()

```

  
#Animation
### Animate India
```{r fig.show='animate'}  
polygon_sources <- 
  india_outline %>%
  #filter(type != "corner") %>% 
  mutate(type = factor(type)) %>% 
  group_by(type) %>%
  nest()

#polygon_sources

polygon_data <- 
  tween_states(polygon_sources$data, tweenlength = 1,
               statelength = 0.5, ease = "sine-out", 
               nframe = 10) %>%
  as.tibble()

polygon_data <-
  polygon_data %>% 
  mutate(.frame = 12 - .frame + 11) %>% 
  rbind(polygon_data)

p <- 
  ggplot(polygon_data, aes(long, lat)) +
  geom_polygon(aes(frame = .frame, group = group), 
               fill = "white", color = "grey70") +
  geom_point(aes(X, Y), data = data_raw, size = 3)


animation::ani.options(interval = c(1/2))
gganimate(p, title_frame = FALSE)
```

#All together
```{r fig.show='animate', fig.height = 7, fig.width = 8}
p <- 
  ggplot(polygon_data, aes(long, lat)) +
  geom_polygon(aes(frame = .frame, group = group), 
               fill = "white", color = "grey80") +
  geom_point(aes(X, Y, color = Improved, frame = .frame), size = 3, 
             data = points_data) +
  geom_path(aes(X, Y, group = State, color = Improved, frame = .frame), 
            data = points_data) +
  guides(color = F, size = F) +
  geom_text(aes(X, Y, label = label, frame = .frame,
                size = size, color = color), 
            points_text, hjust = 0) +
  scale_radius(range = c(5, 8)) +#  geom_text(aes(ifelse(Pct_Usable_Toilets > 2014, X+1, X-1), 
#                ifelse(Pct_Usable_Toilets > .5, Y+1, Y-1), 
#                group = State, label = Label, frame = .frame), 
#                  data = filter(points_data, !is.na(Label))) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12),
        plot.caption = element_text(size = 10)) +
  coord_fixed() +
  labs(title = "Access to working toilets still in need of improvement",
       subtitle = str_wrap("Although access to working toilets in rural elementary schools has grown in most states of India (2010 - 2016), many states stuggle to have at least 60% of toilets available for students. Unfortunately, these stats are worse for girls than for boys.", width = 85), 
        caption = "Jake Riley - #MakeoverMonday Week 32 \npersonal goal: create smooth animations with R")

frame_times <-
  c(rep(1, 2), 4, rep(1/2, 6), 2, rep(1/5, 12))

animation::ani.options(interval = frame_times)

gganimate(p = p, 
          #filename = "india.gif", 
          #saver = "gif",
          ani.width = 600, 
          ani.height = 700,
          title_frame = F)

```

The ggmap library is provided by:
  
  D. Kahle and H. Wickham. ggmap: Spatial Visualization with
ggplot2. The R Journal, 5(1), 144-161. URL
http://journal.r-project.org/archive/2013-1/kahle-wickham.pdf
