---
title: "United Hexagons of Waffles"
author: "Jake Riley" 
date: "February 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F,
                      fig.height = 12, fig.width = 12)
```

This project combines the hexmap and waffle grid data visualization techniques to create a waffle grid that sits witin each hexagon. I will not plot the hexagon in this example but I will show you how to set up the waffle grid for each state.

###Setup workspace
```{r}
setwd("C:\\Users\\HP\\Desktop")
library(plotly)
library(tidyverse)
```

###The original data
I used ```dput()``` to get the structure of my original data. The ```Type``` and ```Tile``` columns are so that I can union this data (hex_map) with the dataset in the next step (hex_grid)
```{r}
#I used the dput formula to give you a reproducible example
hex_map <-
  structure(
    list(
        Row = 
          c(8, 2, 3, 3, 3, 4, 5, 3, 4, 0, 1, 0, 5, 5, 5, 5, 3, 4, 2, 6, 4, 8, 6, 6, 4, 
            2, 6, 3, 6, 4, 7, 5, 2, 4, 6, 5, 2, 4, 5, 6, 2, 5, 3, 1, 3, 3, 7, 6, 6, 4, 5), 
        Column = 
          c(0.5, 7.5, 6, 3, 2, 3.5, 11, 12, 9.5, 8.5, 8, 0.5, 5, 2, 6, 7, 5, 6.5, 5.5, 
            10.5, 8.5, 11.5, 7.5, 4.5, 5.5, 6.5, 2.5, 9, 3.5, 4.5, 11, 10, 3.5, 2.5, 
            9.5, 8, 4.5, 1.5, 9, 11.5, 8.5, 4, 7, 4, 4, 8, 10, 1.5, 5.5, 7.5, 3), 
        State = 
          c("Alaska", "Alabama", "Arkansas", "Arizona", "California", "Colorado", 
            "Connecticut", "District of Columbia", "Delaware", "Florida", "Georgia", 
            "Hawaii", "Iowa", "Idaho", "Illinois", "Indiana", "Kansas", "Kentucky", 
            "Louisiana", "Massachusetts", "Maryland", "Maine", "Michigan", "Minnesota", 
            "Missouri", "Mississippi", "Montana", "North Carolina", "North Dakota", 
            "Nebraska", "New Hampshire", "New Jersey", "New Mexico", "Nevada", 
            "New York", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", 
            "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Virginia",
            "Vermont", "Washington", "Wisconsin", "West Virginia",
            "Wyoming"), 
        Abbreviation = 
          c("AK", "AL", "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA", "HI", 
            "IA","ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME", "MI", "MN", 
            "MO", "MS", "MT", "NC", "ND", "NE", "NH", "NJ", "NM", "NV", "NY", "OH", 
            "OK", "OR", "PA","RI", "SC", "SD", "TN", "TX", "UT", "VA", "VT", "WA", 
            "WI", "WV", "WY"),
        Type = rep("Hex", 51), 
        Tile = rep(0, 51)
    ), 
    class = "data.frame", 
    .Names = c("Row", "Column", "State", "Abbreviation", "Type", "Tile"), 
    row.names = c(NA, -51L))
```

###Here's what it looks like:
I created this canvas of overlapping circles to simulate a hex grid
```{r echo = F}
test_canvas <-
  #create a XY matrix
    expand.grid(X = c(-0:13),
                Y = c(-1:9),
                KEEP.OUT.ATTRS = F) %>% 
  #offset the circles of the even rows
    mutate(X = ifelse(Y %% 2 == 0, X - 0.5, X)) 

```

And then lay my states on top of it
```{r}
gg_canvas <-
  ggplot() +
  geom_point(data = test_canvas, aes(X, Y), size = 28, alpha = .03)+
  geom_point(data = hex_map, aes(Column, Row), size = 28, alpha = .2)+
  geom_text(data = hex_map, aes(Column, Row + 0.5, label = Abbreviation), 
             size = 4, fontface = "bold")+
  coord_fixed(ratio = 1) +
  xlim(-2, 14) +  ylim(-2, 14) +
  theme_void()

gg_canvas
```

###Parameters
Before I create the waffle grid data frame. I need to set up some parameters:
*```grid_modulo``` controls when to switch to a new row
*```grid_width```  controls the distance of the waffle chart
*```grid_offset``` repositions the bottom left corner of the grid based on the centroid of the hexagon
*```grid_spacing``` use the 

```{r}
#number of columns for grid. 1, 11, 21 in the first column
  grid_modulo <- 10

#the distance between centroids of the hexagons is 1 so I want to create a distance .7 wide. To do this, I divide 'grid_width' by 2 and move the starting point 0.35 down and 0.35 over with 'grid_offset'
  grid_width <- .7
  grid_offset <- grid_width/2   
  
#once the columns are established the column number can be multiplied by the 1 tenth of the total width. Column 1 will start on 1*.07, Column 2 will start on 2*.07, etc
  grid_spacing <- grid_width/grid_modulo

#create a matrix to get the X & Y coordinates of the tiles
  waffle_matrix <-
    matrix(data = c(1:100), nrow = 10) %>% 
    t()
  
  waffle_matrix  

#this formula finds the position of '2' in the matrix
  which(waffle_matrix == 2, arr.ind = T)

#We can use this to create functions
  return_row <-
    function(x){
      which(waffle_matrix == x, arr.ind = T)[1] #row = 1st value
      #((x - 1) %/% grid_modulo) #using modular arithmetic
    }

  return_col <- 
    function(x){
      which(waffle_matrix == x, arr.ind = T)[2] #column = 2nd value
      #(x-((x-1) %/% grid_modulo)*grid_modulo) #using modular arithmetic
    }

#make sure they work. Expand the console to fit the 10 values per row
  data.frame(Tile = c(1:20)) %>% 
  rowwise() %>% 
  mutate(row = return_row(Tile),
         col = return_col(Tile))  %>% 
  as.data.frame()

```

```{r fig.height = 4, fig.width = 4}
#build the table. First, 
hex_grid <-
  #fist create tiles 1:100 for each state (hex_map$Abbreviation)
    expand.grid(Abbreviation = hex_map$Abbreviation,
              Tile = as.numeric(c(1:100)),
              KEEP.OUT.ATTRS = F,
              stringsAsFactors = F) %>% 
    arrange(Abbreviation) %>% 
  #join in the original Row & Column coordinates
    left_join(select(hex_map, Row:Abbreviation)) %>%
    rowwise() %>% 
    mutate(
       Type = "Waffle",
     #first move the starting point of the Row & Columns
       AdjustCol = Column - grid_offset,
       AdjustRow = Row - grid_offset,
     #use the functions above on the value of [Tile] to set new coordinates
       WaffleCol = return_col(Tile) - .5,
       WaffleRow = return_row(Tile) - .5,
     #use the original column + (the waffle column * grid spacing)
       FinalCol = AdjustCol + (WaffleCol * grid_spacing),
       FinalRow = AdjustRow + (WaffleRow * grid_spacing)
     )

#confirm that grid is iin the correct order & aligned with the centroid
ggplot(filter(hex_grid, Abbreviation == "PA"), aes(x = Column, y = Row)) +
  geom_point(size = 115, shape = 4, stroke = 3, color = "grey60") +
  geom_point(size = 115, shape = 0, stroke = 3, color = "grey60") +
  geom_point(size = 5) +
  geom_label(aes(x = FinalCol, y = FinalRow, label = Tile, fill = Tile), 
             alpha = .8, show.legend = F) +
  scale_fill_continuous(low = "yellow", high = "orange") +
  coord_equal() +
  theme_void()
```

A sample of the map
```{r}
gg_canvas +
  geom_point(data = hex_grid,
             aes(x = FinalCol, y = FinalRow, color = Tile < 37), 
             show.legend = F, size = .9) + #theme_void() +
  scale_color_manual(values = c("white", "#00C0AF"))
```

###The final data frame
```{r}
final_hex <-
  #take some columns from the new data frame and then rename them
    hex_grid %>% 
    select(State, Abbreviation, Tile, Type, FinalCol, FinalRow) %>% 
    rename(Column = FinalCol,
           Row = FinalRow) %>% 
  #union the original dataframe to the new one
    rbind(hex_map)

#write dataframe
  write.csv(final_hex, "final_hex.csv", row.names = F)
```